workshop

### 2022-09-12 scripts/workshop.Rmd
cc and bc distances

ggplot(bc_dist_df, aes(x = col, y = row, fill = bc_dist)) + geom_tile()
```

Manhattan method is the absolute value of the difference between two values
So higher distance means larger difference
Normalized to have range 0 to 1
```{r}
cc_diff <- comm_summ %>% dplyr::select(siteID, cc) %>% mutate(log_cc = log1p(cc)) %>% dplyr::select(-cc)
cc_dist <- dist(cc_diff, method = "manhattan")
cc_diff_mat <- as.matrix(cc_dist, labels = T)
colnames(cc_diff_mat) <- cc_diff[[1]]
rownames(cc_diff_mat) <- cc_diff[[1]]
cc_diff_df <- melt(cc_diff_mat, varnames=c("row", "col"), value.name = "cc_dist") %>% 
  mutate(cc_norm = (cc_dist-min(.$cc_dist))/(max(.$cc_dist)-min(.$cc_dist)))

ggplot(cc_diff_df, aes(x = col, y = row, fill = cc_norm)) + geom_tile()
```

```{r}
bc_cc_df <- left_join(cc_diff_df, bc_dist_df)
bc_cc_df %<>% mutate(diff = abs(bc_dist-cc_norm))

@@

bc_cc_df %>% filter(bc_dist!=1) %>% ggplot(.,aes(x=cc_norm, y=bc_dist)) + geom_point() + geom_smooth(method = "lm")
```

>>>>

bc_dist_df %>% filter(bc_dist!=0) %>% ggplot(., aes(x=bc_dist)) + geom_histogram()
ggplot(bc_dist_df, aes(x = col, y = row, fill = bc_dist)) + geom_tile()
```

quick check on normality of community competence data
```{r}
comm_summ %>% ggplot(.,aes(x=cc)) + geom_histogram()
```

Manhattan method is the absolute value of the difference between two values
So higher distance means larger difference
Normalized distances to have range 0 to 1
```{r}
cc_diff <- comm_summ %>% dplyr::select(siteID, cc) %>% mutate(log_cc = log1p(cc)) %>% dplyr::select(-cc)
cc_dist <- dist(cc_diff, method = "manhattan")
cc_diff_mat <- as.matrix(cc_dist, labels = T)
colnames(cc_diff_mat) <- cc_diff[[1]]
rownames(cc_diff_mat) <- cc_diff[[1]]
cc_diff_df <- melt(cc_diff_mat, varnames=c("row", "col"), value.name = "cc_dist") %>% 
  mutate(cc_norm = (cc_dist-min(.$cc_dist))/(max(.$cc_dist)-min(.$cc_dist)))

cc_diff_df %>% filter(cc_norm != 0) %>% ggplot(., aes(x=cc_norm)) + geom_histogram()
ggplot(cc_diff_df, aes(x = col, y = row, fill = cc_norm)) + geom_tile()
```

```{r}
bc_cc_df <- left_join(cc_diff_df, bc_dist_df)
```

Each point is a pairwise interaction

The null hypothesis is that community competence is achieved through specific community compositions. i.e. you cannot achieve the same community competence with different community compositions

So if a pair of sites have very different community competence then they should also have very different community composition

But if community competence can be achieved through a variety of different community compositions then a pair of sites with similar community competence can have a high level of dissimilarity

So we are looking for low cc distance between two sites but high bc distance
An upward trend is expected because more dissimilar sites in terms of competence probably have different composition
```{r}
bc_cc_df %>% filter(diff!=0) %>% ggplot(.,aes(x=cc_norm, y=bc_dist)) + geom_point() + geom_smooth(method = "lm")
```

If diff is set as community competence similarity minus community composition similarity then a low value will mean the difference between composition similarity is large compared to competence similarity. Either competence similarity is very high (cc dist is low) and composition similarity (bc dist) is moderate or high OR composition similarity is low (i.e. BC dist is high) and competence similarity (cc dist is moderate or low)

```{r}
bc_cc_df %<>% mutate(diff = cc_norm-bc_dist)

@@

bc_cc_df %>% filter(diff!=0) %>% ggplot(.,aes(x=diff)) + geom_histogram()
```
